Statically Typed String Sanitation Inside a Python (Technical Report)

### 2014-12-14 sanitation-psp14/tr/strings_chapter.tex
diagrams.
nop.
Added example of new case for => direction of semantic alignment

@@

% Application case.
\item \textbf{Case $e=e_1(e_2)$}.
The only applicable rule is S-E-App. By inversion, we establish that the following:
\begin{align*}
  e_1 &\sreduces \lambda x.e_1' \\
  e_2 &\sreduces v_2 \\
  [v_2/x]e_1' &\sreduces v
\end{align*}

@@

@@

\section{String Substitution and Language Replacement}

\subsection{The Trivial Definition}

\subsection{An Automaton Construction}

Insert Automaton stuff...

\subsection{Toward a Precise Definition}

@@

<$e$> ::= 
      $x$ | $\lambda x . e$ | $e(e) $\hfill source terms \alt 
      $\strin{s}$ | $\rsconcat{e}{e}$ | $\strcase{e}{e}{x,y.e}$ \hfill $s \in \Sigma^{*}$ \alt
      $\sisubst{r}{e}{e}$ | $\rcoerce{r}{e}$ | $\rcheck{r}{e}{x.e}{e}$

@@

<$\iota$> ::= $x$ | $\lambda x . \iota$ | $\iota(\iota)$ \hfill target terms \alt
$\tstr{s}$ | $\tconcat{\iota}{\iota}$ | $\pstrcase{\iota}{\iota}{x,y.\iota}$ \alt
  $\rx{r}$ | $\preplace{\iota}{\iota}{\iota}$ | $\tcheck{\iota}{\iota}{\iota}{\iota}$

  <$\dot{v}$> ::= $\lambda x . \iota$ | \tstr{s} | \rx{r} \hfill target values

@@

\inferrule[S-T-Replace]
{\sctx \vdash e_1 : \stringin{r_1} \\ \sctx \vdash e_2 : \stringin{r_2} \\\\ \lreplace{r}{r_1}{r_2} = r'}
{\sctx \vdash \sisubst{r}{e_1}{e_2} : \stringin{r'}}

@@

\begin{figure}
  \small
$\fbox{\inferrule{}{e \ \val}}$

\begin{mathpar}
\inferrule[L-Val]
{  }
{\lambda x:\tau.t \ \val}
\end{mathpar}

$\fbox{\inferrule{}{e \mapsto e}}$

@@

\inferrule[L-E-AppLeft]
{e_1 \mapsto e_1'}
{e_1 e_2 \mapsto e_1' e_2}

\inferrule[L-E-AppRight]
{e_2 \mapsto e_2'}
{e_1 e_2 \mapsto e_1 e_2'}

@@

\inferrule[RT-Trans]
{ e \mapsto^* e' \\ e' \mapsto^* e'' }
{ e \mapsto^* e'' }

\inferrule[RT-Step\footnote{This is a special case of RT-Trans which we distinguish in order to write clearer proofs.}]
{ e \mapsto^* e' \\ e' \mapsto v }
{ e \mapsto^* v }

@@

\begin{figure}
  \small
%$\fbox{\inferrule{}{e \mapsto e}} \  \fbox{ \inferrule{}{v \ \val}}$
\begin{mathpar}
\inferrule[SS-E-RStr]
{ \ }
{\strin{s} \ \val}

\inferrule[SS-E-Concat-Left]
{e_1 \mapsto e_1'}
{\rsconcat{e_1}{e_2} \mapsto \rsconcat{e_1'}{e_2}}

\inferrule[SS-E-Concat-Right]
{e_2 \mapsto e_2'}
{\rsconcat{e_1}{e_2} \mapsto \rsconcat{e_1}{e_2'}}

@@

\inferrule[SS-E-Case-Right]
{
  e_2 \mapsto e_2'
}
{
  \strcase{e_1}{e_2}{x,y.e_3} \mapsto \strcase{e_1}{e_2'}{x,y.e_3}
}

@@

\inferrule[SS-E-Replace-Left]
{e_1 \mapsto e_1'}
{ \sisubst{r}{e_1}{e_2} \mapsto \sisubst{r}{e_1'}{e_2}}

@@

\inferrule[SS-E-Check-StepRight]
{e_2 \mapsto e_2'}
{\rcheck{r}{e}{x.e_1}{e_2} \mapsto \rcheck{r}{e}{x.e_1}{e_2'}}

\inferrule[SS-E-Check-Ok]
{s \in \lang{r}}
{\rcheck{r}{\strin{s}}{x.e_1}{e_2} \mapsto [\rstr{s} / x]e_1}
\inferrule[SS-E-Check-NotOk]
{s \not \in \lang{r}}
{\rcheck{r}{\strin{s}}{x.e_1}{e_2} \mapsto e_2}

@@

\inferrule[P-E-Replace]
{\iota_1 \treduces \rx{r} \\ \iota_2 \treduces \tstr{s_2} \\ \iota_3 \treduces \tstr{s_3} \\ \lsubst{r}{s_2}{s_3} = s} 
{\preplace{\iota_1}{\iota_2}{\iota_3} \treduces \tstr{s}}

\inferrule[P-E-Check-OK]
{
  \iota_r \treduces \rx{r} \\
  \iota \treduces \tstr{s} \\ 
s \in \lang{r} \\ 
\iota_1 \treduces \dot{v_1} 
}
{
  \tcheck{\iota_r}{\iota}{\iota_1}{\iota_2} \treduces \dot{v_1}
}

\inferrule[P-E-Check-NotOK]
{\iota_r \treduces \rx{r} \\ \iota \treduces \tstr{s} \\ s \not \in \lang{r} \\ \iota_2 \treduces \dot{v_2}}
{\tcheck{\iota_r}{\iota}{\iota_1}{\iota_2} \treduces \dot{v_2}}
\end{mathpar}
\caption{Big step semantics for $\lambdap$}.
\label{fig:dlambdap}
\end{figure}

>>>>

% bookmark1
\begin{trlem} \label{lem:inductive}
If $e \sreduces v$ and $e \mapsto e'$ then $e' \sreduces v$. 
\end{trlem}
\begin{proof}
By induction on the structure of e.
\begin{itemize}[label=$ $,itemsep=1ex]

\item \textbf{Case $e = e_1(e_2)$}.
The only applicable rule is S-E-App, so $e_1 \sreduces \lambda x . e_3$
and $e_2 \sreduces v_2$ such that $[v_2/x] e_3 \sreduces v$.

The term $e = e_1(e_2)$ may step by three rules.

First, if $e$ steps by L-E-AppLeft then $e_1(e_2) \mapsto e_1'(e_2)$.
By induction, $e_1' \sreduces \lambda x . e_3$.
By S-E-App, $e_1'(e_2) \sreduces \lambda x . e_3$.

Second, if $e$ steps by L-E-AppRight then $e_1(e_2) \mapsto e_1(e_2')$.
By induction, $e_2' \sreduces v_2$.
By S-E-App, $e_1(e_2') \sreduces \lambda x . e_3$.

Third, if $e$ steps by L-E-AppAbs then $e = (\lambda x . e_3)(v_2) \mapsto [v_2 / x]e_3$.
By induction, $e = (\lambda x . e')v_2 \sreduces v$.
\end{itemize}
\end{proof}

@@

% bookmark2
% Application case.
\item \textbf{Case $e=e_1(e_2)$}.
The only applicable rule is S-E-App. By inversion:
\begin{align*}
  e_1 &\sreduces \lambda x.e_1' \\
  e_2 &\sreduces v_2 \\
  [v_2/x]e_1' &\sreduces v
\end{align*}

@@

If $e_1 = \lambda x . e_1'$ and $e_2 = v_2$ (henceforth the reflexive case)
then $e \mapsto [v_2 / x] e_1'$ and the conclusion follows by RT-Trans.

If $e_1 \mapsto \lambda x . e_1'$ then $e_1(e_2) \mapsto (\lambda x . e_1')(e_2) \mapsto [e_2/x]e_1'$ and the conclusion follows by two applications
of RT-Trans.

If $e_1 \mapsto^k \lambda x . e_2'$ then $e_1 \mapsto e'$ so $e_1(e_2) \mapsto e'(e_2)$ and by Lemma~\ref{lem:inductive} $e'(e_2) \mapsto^* v$.
So $e_1(e_2) \mapsto e'(e_2) \mapsto^* v$; it follows by RT-Trans that $e_1(e_2) \mapsto v$.

\todo{not modified below this line...}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

@@

<$e$> ::= 
      $x$ | $v$ \hfill source terms \alt 
      $\rsconcat{e}{e}$ | $\strcase{e}{e}{x,y.e}$ \hfill $s \in \Sigma^{*}$ \alt
      $\sisubst{r}{e}{e}$ | $\rcoerce{r}{e}$ | $\rcheck{r}{e}{x.e}{e}$

@@

<$\iota$> ::= $x$ | $\dot{v}$ \hfill target terms \alt
$\tconcat{\iota}{\iota}$ | $\pstrcase{\iota}{\iota}{x,y.\iota}$ \alt
  $\rx{r}$ | $\preplace{\iota}{\iota}{\iota}$ | $\tcheck{\iota}{\iota}{\iota}{\iota}$

  <$\dot{v}$> ::= $\lambda x . \iota$ | $\tstr{s}$ | $\rx{r}$ \hfill target values

@@

\inferrule[S-T-Replace]
{\sctx \vdash e_1 : \stringin{r_1} \\ \sctx \vdash e_2 : \stringin{r_2} }
{\sctx \vdash \sisubst{r}{e_1}{e_2} : \stringin{\lreplace{r}{r_1}{r_2}}}

@@

\begin{figure}
  \small
$\fbox{\inferrule{}{e \mapsto e}}$

@@

\inferrule[L-E-AppLeft]
{e_1 \mapsto e_1'}
{e_1(e_2) \mapsto e_1'(e_2)}

\inferrule[L-E-AppRight]
{e_2 \mapsto e_2'}
{v_1 \mapsto v_1}

@@

\inferrule[RT-Trans]
{ e \mapsto^* e' \\ e' \mapsto e'' }
{ e \mapsto^* e'' }

@@

\begin{figure}
  \small
$\fbox{\inferrule{}{e \mapsto e}}$ (Contiues figure 6)
%$\fbox{\inferrule{}{e \mapsto e}} \  \fbox{ \inferrule{}{v \ \val}}$
\begin{mathpar}
\inferrule[SS-E-Concat-Left]
{e_1 \mapsto e_1'}
{\rsconcat{e_1}{e_2} \mapsto \rsconcat{e_1'}{e_2}}

\inferrule[SS-E-Concat-Right]
{e_2 \mapsto e_2'}
{\rsconcat{v_1}{e_2} \mapsto \rsconcat{v_1}{e_2'}}

@@

@@

\inferrule[SS-E-Replace-Left]
{e_1 \mapsto e_1'}
{ \sisubst{r}{v_1}{e_2} \mapsto \sisubst{r}{v_1'}{e_2}}

@@

\inferrule[SS-E-Check-Ok]
{s \in \lang{r}}
{\rcheck{r}{\strin{s}}{x.e_1}{e_2} \mapsto [\rstr{s} / x]e_1}

\inferrule[SS-E-Check-NotOk]
{s \not \in \lang{r}}
{\rcheck{r}{\strin{s}}{x.e_1}{e_2} \mapsto e_2}

@@

\inferrule[P-E-Replace]
{\iota_1 \treduces \rx{r} \\ \iota_2 \treduces \tstr{s_2} \\ \iota_3 \treduces \tstr{s_3} } 
{\preplace{\iota_1}{\iota_2}{\iota_3} \treduces \tstr{\lsubst{r}{s_2}{s_3} }}

\inferrule[P-E-Check-OK]
{
  \iota_x \treduces \rx{r} \\
  \iota \treduces \tstr{s} \\ 
s \in \lang{r} \\ 
\iota_1 \treduces \dot{v_1} 
}
{
  \tcheck{\iota_x}{\iota}{\iota_1}{\iota_2} \treduces \dot{v_1}
}

\inferrule[P-E-Check-NotOK]
{\iota_x \treduces \rx{r} \\ \iota \treduces \tstr{s} \\ s \not \in \lang{r} \\ \iota_2 \treduces \dot{v_2}}
{\tcheck{\iota_x}{\iota}{\iota_1}{\iota_2} \treduces \dot{v_2}}
\end{mathpar}
\caption{Big step semantics for $\lambdap$}.
\label{fig:dlambdap}
\end{figure}
